---
title: "Data Analysis on the nations GDP"
author: "Triaccau"
date: "26/6/2021"
output: 
 prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## Highest nations GDP in 2018


First, it might be interesting to look at the ranking of the top 10 nations with the highest GDP. It is not a surprise that China, USA and India are at the top of this specific ranking. Asia is the most represented continent on this list, with 5 nations. Europe is in second place, thanks to France, Germany and the United Kingdom. Instead, no nation representing the African continent.


```{r}
library(dplyr)

library(ggplot2)

library(forcats)

library(prettydoc)

Debt_nations <- read.csv("total_gov_expenditure_gdp_data.csv")

GDP_nations <- read.csv("gdp_world_regions_stacked_data.csv")

top_10_GDP <- GDP_nations %>%
select(Entity, Code, Year, GDP) %>%
filter(Year == 2018 & (Code != "" & Code != "OWID_WRL")) %>%
group_by(Entity) %>%
arrange(desc(GDP)) %>%
head(., 10)

top_10_GDP



top_10_GDP$Entity <- as.factor(top_10_GDP$Entity)

ggplot(top_10_GDP, aes(GDP, fct_reorder(Entity, GDP))) + geom_col() + labs(title = "World GDP Countries", subtitle = "Top 10 countries with highest GDP", y = "Country")

```

## Regions trend among years (1950 - 2018)

Another interesting aspect is the growth of GDP by different region since 1950 to 2018. In this case is possible to observe the huge increasing of home production in the east Asia region, western Europe, south and south east Asia and western offshoots (eg USA, Canada, Australia).

```{r}
GDP_trend_byregion <- GDP_nations %>%
filter(Code == "" & Year >= 1950) %>%
arrange(desc(Entity))

ggplot(GDP_trend_byregion, aes(Year, GDP, color = Entity)) + geom_line() + labs(title = "GDP trend", subtitle = "Growing of GDP along regions between 1850 and 2018")
```

## Highest nations GDP growth along years

At the end, we can analyze the nations with the hugest increasing percent level of GDP during the period 1950 - 2018. The ranking is dominated by middle east countries. As it is possible to see UAE, Oman and Qatar are on the top; naturally, the discovery of petrol in this territories represent the main motive of these great rates of increasing.

```{r}
GDP_2018 <- GDP_nations %>%
group_by(Entity) %>%
filter(Year == "2018" & Code != "" & Code != "OWID_WRL") %>%
arrange(Entity) %>%
select(Entity, GDP)

GDP_1950 <- GDP_nations %>%
group_by(Entity) %>%
filter(Year == "1950" & Code != "" & Code != "OWID_WRL") %>%
arrange(Entity) %>%
select(Entity, GDP)

top_10_GDP_grow_rate <- GDP_2018 %>%
inner_join(GDP_1950, by = "Entity", suffix = c(".2018", ".1950")) %>%
mutate(grow_rate = (GDP.2018 - GDP.1950) / GDP.1950) %>%
arrange(desc(grow_rate)) %>%
head(., 10)

top_10_GDP_grow_rate$Entity <- as.factor(top_10_GDP_grow_rate$Entity)

ggplot(top_10_GDP_grow_rate, aes(grow_rate, fct_reorder(Entity, grow_rate))) + geom_col() + labs(title = "Highest grow rates", subtitle = "Ten countries with highest grow rate amon 1950 and 2018", y = "Country")
```

